{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-01-28 swmp_noise.tex
minor edits to revision comment
Final revisions for STICS

Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.  Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.  Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.  Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions of ecosystem production can arise from other habitats, such as the benthos or seagrass patches.  By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}. Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.  Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches. Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).  Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in mean annual values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed.

>>>>

Ecosystem metabolism describes the balance between production and respiration processes that create and consume organic matter \citep{Kemp12,Needoba12}.  Light exposure experiments of water samples collected at discrete locations and times have traditionally been used to measure metabolic activity.  Although highly controlled and precise, bottle-based methods are labor-intensive and not scalable to describe entire ecosystem rates.  Bottle-based methods may also only reliably estimate production and respiration associated with planktonic processes, whereas significant contributions to ecosystem metabolism can arise from other habitats, such as the benthos or intertidal marsh areas.  By contrast, open-water techniques have been increasingly used to estimate whole system metabolism given the availability of long-term, continuous time series of dissolved oxygen \citep{Odum56,Davanzo96}. Daily integrated measurements of metabolism represent the balance between daytime production and nighttime respiration attributed to all ecosystem components.  Open-water estimates also provide a basis for tracking ecosystem change over time and are more practical for capturing events or evaluating trends as compared to bottle-based approaches. Although metabolic rates vary naturally at different spatial and temporal scales \citep{Ziegler98,Caffrey04,Russell07}, anthropogenic nutrient sources are often contributing factors that increase rates of production (\citealt{Nixon95}, \citetalias{NRC00}).  Reliables estimates of whole ecosystem metabolism are critical for measuring both background rates of production and potential impacts of human activities on ecosytem condition.

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 9.15 (as a percentage of the total estimates, Rookery Bay) to 21.80 (Padilla Bay) for production and 7.57 (Rookery Bay) to 20.68 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-63.1 \% change from the annual mean) and respiration (-62.6 \%) for Elkhorn Slough, increased mean production (17.8 \%) and respiration (18.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.
